Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
setwd('/Users/Cinkie/Documents/Nanodegree/Data Analysis with R')
library(ggplot2)
pf = read.csv('pseudo_facebook.tsv', sep = '\t')

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() + 
    stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
    geom_line(aes(color = gender), stat = 'summary', fun.y = median) 
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
library(dplyr)
pf.fc_by_age_gender = pf %>%
    filter(!is.na(gender)) %>%
    group_by(age, gender) %>%
    summarise(mean_friend_count = mean(friend_count),
              median_friend_count = median(as.numeric(friend_count)),
              n = n()) %>%
    ungroup() %>%
    arrange(age)


ggplot(pf.fc_by_age_gender, aes(x=age, y=median_friend_count)) + 
    geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
library(reshape2)

pf.fc_by_age_gender.wide = dcast(pf.fc_by_age_gender,
                                 age ~ gender,
                                 value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(pf.fc_by_age_gender.wide, aes(x=age, y=female/male)) + geom_line() + 
    geom_hline(yintercept=1, color = 'red', linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined = floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket = cut(pf$year_joined, c(2004,2009,2011,2012,2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(pf, aes(x=age,y=friend_count)) +
    geom_line(aes(color=year_joined.bucket), stat='summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(subset(pf, !is.na(pf$year_joined.bucket)), aes(x=age,y=friend_count)) +
    geom_line(aes(color=year_joined.bucket), stat='summary', fun.y = mean) + 
    geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
friend_rate = subset(pf, tenure>0)$friend_count / subset(pf, tenure>0)$tenure
summary(friend_rate)

with(subset(pf, tenure > 0), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.2205

What is the maximum friend rate? 0.6096

```{r Friendships Initiated}
ggplot(subset(pf,tenure>=1), aes(x=tenure, y=friendships_initiated/tenure)) +
    geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

p1=ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2=ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3=ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4=ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

grid.arrange(p1,p2,p3,p4)

p5=ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

grid.arrange(p1,p5)
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo = read.csv('yogurt.csv')
str(yo)

# Convert id from int to factor
yo$id = factor(yo$id)
str(yo)

ggplot(yo, aes(price)) + geom_histogram(binwidth = 1)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo$price)
table(yo$price)

yo = transform(yo, all.purchases = strawberry+blueberry+pina.colada+plain+mixed.berry)
```

***

### Prices over Time
Notes: Prices are steadily increasing.

```{r Prices over Time}
ggplot(yo, aes(x=time, y=price)) + 
    geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
set.seed(0)
sample.ids = sample(levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(x=time, y=price)) + 
    facet_wrap(~id) + 
    geom_line() +
    geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

```{r}
install.packages('GGally')
require(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset = pf[,c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])

with(pf.sample, cor.test(friendships_initiated,friend_count))
with(pf.sample, cor.test(age, mobile_likes))
with(pf.sample, cor.test(
```


### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

